#SQL
***
##Filtering Rows and Columns

In SQL you can filter columns by using SELECT and rows by using WHERE. Let's look at an example using the Lahman baseball database. First we should load the Lahman package. We will also need the sqldf package.

```{r,echo=TRUE,warning=FALSE,message=FALSE}
library(Lahman)
library(sqldf)
```

Let's take a look at an example. Suppose we want to see the homerun totals for the 1927 Yankees. We could write:

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE teamID='NYA'and yearID=1927"

sqldf(query)
```
***
###More Examples of SQL

If one wants to find the specific codes of words for Batting they would use 'colnames(Batting)'. In these examples, SQL is being used. First one needs to load these packages:

```{r}
library(Lahman)

library(sqldf)
```

Example 2: Find all instances where the Yankees have hit 40 or more homeruns.

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting WHERE teamID='NYA' and HR>=40"


sqldf(query)
```

Example 3: Find all instances where a Yankees player had more than 40 homeruns and fewer than 60 strikeouts.

```{r}
query<-"SELECT playerID,teamID,yearID,HR,SO FROM Batting WHERE HR>40 and SO<60"


sqldf(query)
```

Example 4: Find all instances of Phillies players hitting more than 30 homeruns in the 1970s.

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting WHERE teamID='PHI' and yearID>=1970 and yearID<=1979 and HR>30"


sqldf(query)
```
***
##Order By

With SQL, one can order data in descending or ascending order. For example, to find players with more than 50 homeruns from most to least, one would write:

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE HR>50 ORDER BY HR DESC"
sqldf(query)
```
***
###More Examples of Order By

Example 2: Find all instances of a player with fewer than 10 strikeouts, among those players with at least 400 at-bats. Order results from least strikeouts to most.

```{r}
query<-"SELECT playerID,teamID,yearID,SO,AB FROM Batting WHERE SO<10 and AB>=400 ORDER BY SO"
sqldf(query)
```
***
##Aggregation

One may aggregate data by using commands such as sum, min, avg, and max to organize data however necessary. If one wanted to find the career homerun totals for Babe Ruth, it could be done the following way:

```{r}
query<-"select playerID, sum(HR) from Batting where playerID='ruthba01'
group by playerID"
sqldf(query)
```
***
###More Examples of Aggregation

Example 2: Show players with 600 or more career homeruns in descending order:

```{r}
query<-"select playerID, sum(HR) from Batting 
group by playerID 
having sum(HR)>=600
order by sum(HR) desc"
sqldf(query)
```

Example 3: Find players whose average number of homeruns per season is greater than 30, in descending order:

```{r}
query<-"select playerID, avg(HR) from Batting 
group by playerID 
having avg(HR)>30
order by avg(HR) desc"
sqldf(query)
```
***
##Joins

If two tables need to joined, one can use the inner join command. One must also specify which portions of each table are to be joined. Display player first and last name, team, year, and homeruns for Babe Ruth:

```{r}
query<-"select nameFirst, nameLast, teamID, yearID,HR 
from Batting inner join Master
on Batting.playerID=Master.playerID
where Batting.playerID = 'ruthba01'"

sqldf(query)
```
***
###More Join Examples

Example 2: All players who hit more than 50 homeruns, including first and last names, teams, years and homeruns in descending order:

```{r}
query<-"select nameFirst, nameLast, teamID, yearID, HR 
from Batting inner join Master
on Batting.playerID=Master.playerID
where HR>50
order by HR desc"

sqldf(query)
```

Example 3: Display player ID, team name, year and homeruns for Babe Ruth:

```{r}
query<-"select playerID, name, Batting.yearID, Batting.HR
from Batting inner join Teams
on Batting.teamID=Teams.teamID
and Batting.yearID=Teams.yearID
where playerID='ruthba01'"

sqldf(query)
```

Example 4: Display career homeruns for all players. Limit the display to those that have at least 600 homeruns, with the highest at the top. Include first and last names:

```{r}
query<-"SELECT nameFirst,nameLast,sum(HR)
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
GROUP BY Batting.playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"

sqldf(query)
```

Example 5: List all players with the first name Bob who averaged more than $1 Million over their career:

```{r}
query<-"SELECT nameFirst,nameLast,avg(salary)
FROM Salaries INNER JOIN Master
ON Salaries.playerID=Master.playerID
WHERE nameFirst='Bob'
GROUP BY Salaries.playerID
HAVING avg(salary)>1000000
ORDER BY avg(salary) DESC"

sqldf(query)
```

